---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Cassie Lee"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
bibliography: references.bib
---

```{r}
#| label: load-packages-data

library(tidyverse)
library(here)
library(knitr)
```

::: {.callout-tip icon="false"}
## Github Repo Link

[Cassie Lee's Final Project Memo](https://github.com/stat301-2-2024-winter/final-project-2-cassielee2025.git)
:::

## Analysis Plan

### Data Splitting

The 40,000 observations used for this prediction problem was split using a 75/25 split between training and testing data, stratified by birth weight.

```{r}
#| label: data-split
#| echo: false

load(here("memos/memo2_outputs/split_dimensions.rda"))
split_dimensions %>% 
  kable()
```


### Resampling

The training data was then resampled using vfold cross validation with 5 folds and 3 repeats, so each model/workflow will be trained/fit 15 times for a metric estimate and standard error.


**Model Types**

**Recipes**

**Metrics**

## Two Model Types

**Null/base Model**

**Standard Linear Regression**

## Exploratory Data Analysis

## Progress Summary and Potential Issues
